Galactic Chemical Evolution Models Favor an Extended Type Ia Supernova Delay-Time Distribution

### 2024-02-27 src/tex/ms.tex
[showyourwork] overleaf sync

% user-defined commands
\newcommand{\yes}{\textcolor{green}{\checkmark}}
\newcommand{\meh}{\textcolor{black}{$\sim$}}
\newcommand{\no}{\textcolor{red}{$\times$}}
\newcommand{\osuaffil}{Department of Astronomy, The Ohio State University, 140 W. 18th Ave, Columbus OH 43210, USA}
\newcommand{\ccappaffil}{Center for Cosmology and AstroParticle Physics, The Ohio State University, 191 W. Woodruff Ave., Columbus OH 43210, USA}
\newcommand{\aFe}{[$\alpha$/Fe]\xspace}

@@

% Abstract from AAS 241
\begin{abstract}
    Type Ia supernovae (SNe Ia) produce most of the Fe-peak elements in the Universe and therefore are a crucial ingredient in galactic chemical evolution models. SNe Ia do not explode immediately after star formation and the delay-time distribution (DTD) has not been definitively determined by supernova surveys or theoretical models. Because the DTD also affects the relationship among age, [Fe/H], and \aFe in chemical evolution models, comparison with observations of stars in the Milky Way is an important consistency check for any proposed DTD. We implement several popular forms of the DTD in combination with multiple star formation histories for the Milky Way in one-zone and multi-zone chemical evolution models using the Versatile Integrator for Chemical Evolution (VICE). VICE includes a prescription for radial stellar migration in multi-zone models. We compare our predicted interstellar medium abundance tracks, stellar abundance distributions, and stellar age distributions to the final data release of the Apache Point Observatory Galactic Evolution Experiment (APOGEE). We find that the SN Ia DTD has the largest effect on the \aFe distribution: a DTD with more prompt SNe Ia produces a stellar abundance distribution that is skewed toward a lower \aFe ratio, whereas a DTD with more delayed SNe Ia produces a distribution that is skewed toward higher \aFe. While the DTD alone cannot explain the observed bimodality in the \aFe distribution, in combination with an appropriate star formation history it affects the goodness of fit between the simulated and observed high-$\alpha$ sequence. Our model results favor a DTD with longer delay times and fewer prompt SNe Ia than the fiducial $t^{-1}$ power law.
\end{abstract}

@@

This paper presents the first comprehensive look at the DTD in a multi-zone GCE model that can qualitatively reproduce the observed abundance structure of the Milky Way disk. We evaluate a selection of DTD models from the literature with multiple SFHs and a prescription for radial stellar migration in the Versatile Integrator for Chemical Evolution \citep[VICE;][]{JohnsonWeinberg2020-Starbursts}. In Section \ref{sec:methods}, we present our models for the DTD and SFH, and we describe our chemical evolution models. In Section \ref{sec:onezone-results}, we present the results of our one-zone chemical evolution models. In Section \ref{sec:multizone-results}, we present the results of our multi-zone models and compare to observations. In Section \ref{sec:discussion}, we discuss the implications for the DTD and future surveys. In Section \ref{sec:conclusions}, we summarize our conclusions.

@@

We make use of the multi-zone simulation tools in VICE developed by \citet{Johnson2021-Migration} (hereafter \citetalias{Johnson2021-Migration}). The basic setup of our models follows theirs: the simulated Galaxy is a disk with a radius of 20 kpc and is divided into concentric rings of width $\delta R_{\rm gal}=100$ pc. The gas in each ring is assumed to be instantaneously mixed; while star particles are allowed to migrate radially (see Appendix \ref{app:migration}), there is no exchange of gas between rings. We run our simulations with a time-step size of $\Delta t=10$ Myr up to a maximum time of $t_{\rm max}=13.2$ Gyr. Following \citetalias{Johnson2021-Migration}, we set VICE to form $n=8$ stellar populations per ring per time-step, and we set a maximum star-formation radius of $R_{\rm SF} = 15.5$ kpc, such that $\dot\Sigma_*=0$ for $R_{\rm gal}>R_{\rm SF}$ (stars are allowed to migrate past $R_{\rm SF}$, they are just not born there).

Table \ref{tab:multizone-parameters} summarizes many of the parameters in our multi-zone models.

\begin{deluxetable*}{Cccl}
    \tablecaption{A summary of parameters and their fiducial values for our chemical evolution models. We omit parameters which have been left unchanged from \citetalias{Johnson2021-Migration}; see their Table 1 for details.\label{tab:multizone-parameters}}
    \tablehead{
        \colhead{Quantity} & \colhead{Fiducial Value(s)} & \colhead{Section} & \colhead{Description}
    }
    \startdata
        % Multi-zone simulation parameters
        R_{\rm gal}     & [0, 20] kpc   & \ref{sec:methods} & Galactocentric radius \\
        \delta R_{\rm gal}  & 100 pc    & \ref{sec:methods} & Width of each concentric ring \\
        \Delta R_{\rm gal}  & N/A       & \ref{app:migration} & Change in orbital radius due to stellar migration \\
        p(\Delta R_{\rm gal}|\tau,R_{\rm form}) & Equation \ref{eq:radial-migration}    & \ref{app:migration} & Probability density function of radial migration \\
        z                   & [-3, 3] kpc                & \ref{app:migration} & Distance from Galactic midplane at present day \\
        p(z|\tau,R_{\rm final}) & Equation \ref{eq:sech-pdf}            & \ref{app:migration} & Probability density function of Galactic midplane distance\\
        \Delta t        & 10 Myr    & \ref{sec:methods} & Time-step size \\
        t_{\rm max}     & 13.2 Gyr  & \ref{sec:methods} & Maximum simulation time \\
        n               & 8         & \ref{sec:methods} & Number of stellar populations formed per ring per time-step \\
        R_{\rm SF}      & 15.5 kpc  & \ref{sec:methods} & Maximum radius of star formation \\
        M_{g,0}   & 0         & \ref{sec:sfh}     & Initial gas mass \\
        % DTD
        R_{\rm Ia}(t)   & Equations \ref{eq:powerlaw-dtd}--\ref{eq:prompt-dtd}  & \ref{sec:dtd-models}  & Delay time distribution of Type Ia supernovae \\
        t_D             & 40 Myr    & \ref{sec:dtd-models}  & Minimum SN Ia delay time \\
        N_{\rm Ia}/M_*  & $2.2\times10^{-3}$ M$_\odot^{-1}$ & \ref{sec:yields}  & SNe Ia per unit mass of stars formed \citep{MaozMannucci2012-SNeIaReview} \\
        % Nucleosynthetic yields
        y_{\rm O}^{\rm CC}  & 0.015     & \ref{sec:yields}  & CCSN yield of O    \\
        y_{\rm Fe}^{\rm CC} & 0.0012    & \ref{sec:yields}  & CCSN yield of Fe   \\
        y_{\rm O}^{\rm Ia}  & 0         & \ref{sec:yields}  & SN Ia yield of O       \\
        y_{\rm Fe}^{\rm Ia} & 0.00214   & \ref{sec:yields}  & SN Ia yield of Fe \\
        % Star formation histories
        f_{\rm IO}(t|R_{\rm gal})   & Equation \ref{eq:insideout-sfh}   & \ref{sec:sfh} & Time-dependence of the SFR in the inside-out model \\
        f_{\rm LB}(t|R_{\rm gal})   & Equation \ref{eq:lateburst-sfh}   & \ref{sec:sfh} & Time-dependence of the SFR in the late-burst model \\
        \uptau_{\rm EB}(t)          & Equation \ref{eq:earlyburst-taustar}  & \ref{sec:sfh}   & Time-dependence of the SFE timescale in the early-burst model \\
        f_{\rm EB}(t|R_{\rm gal})   & Equation \ref{eq:earlyburst-ifr}  & \ref{sec:sfh} & Time-dependence of the IFR in the early-burst model \\
        f_{\rm TI}(t|R_{\rm gal})   & Equation \ref{eq:twoinfall-ifr}   & \ref{sec:sfh} & Time-dependence of the IFR in the two-infall model \\
    \enddata
\end{deluxetable*}

@@

Like \citetalias{Johnson2021-Migration}, we focus our analysis on O and Fe, representing the $\alpha$ and Fe-peak elements, respectively. Both elements are produced by CCSNe. VICE adopts the instantaneous recycling approximation for CCSNe, so the equation which governs CCSN enrichment as a function of star formation for some element $x$ is simply
\begin{equation}
    \dot M_x^{\rm CC} = y_x^{\rm CC} \dot M_*
    \label{eq:ccsn-enrichment}
\end{equation}
where $y_x^{\rm CC}$ is the CCSN yield of element $x$ per unit mass of star formation, and $\dot M_*$ is the SFR. Following \citetalias{Johnson2021-Migration}, who in turn adopt their yields from \citet{Weinberg2017-ChemicalEquilibrium}, we adopt $y_{\rm O}^{\rm CC}=0.015$ and $y_{\rm Fe}^{\rm CC}=0.0012$. The primary effect of these yields is to set the ``plateau'' in [O/Fe] which represents pure CCSN enrichment. The chosen yields for this paper produce a plateau at [O/Fe]$=0.45$; see \citet{Weinberg2023-CCSNYield} for more discussion.

@@

The yield $y_{\rm Fe}^{\rm Ia}$ measures the mass of Fe produced by SNe Ia per unit mass of star formation. It can be expressed in terms of the integral of the DTD:
\begin{equation}
    y_{\rm Fe}^{\rm Ia} = m_{\rm Fe}^{\rm Ia} \int_{t_D}^{t_{\rm max}} R_{\rm Ia}(t')dt' = m_{\rm Fe}^{\rm Ia} \frac{N_{\rm Ia}}{M_*}
    \label{eq:dtd-integral}
\end{equation}
where $m_{\rm Fe}^{\rm Ia}$ is the average mass of Fe produced by a single SN Ia. We take the average number of SNe Ia per mass of stars formed to be $N_{\rm Ia}/M_*=2.2\times10^{-3}\,{\rm M}_\odot^{-1}$ from \citet{MaozMannucci2012-SNeIaReview}.%, which gives $R_{\rm Ia}$ units of ${\rm M}_\odot^{-1}\,{\rm yr}^{-1}$. 
Following \citetalias{Johnson2021-Migration}, we take $y_{\rm Fe}^{\rm Ia}=0.00214$. This yield was selected so that the inside-out SFH model produces stars with [O/Fe]$\approx 0.0$ by the end of the simulation. SNe Ia do not contribute O so $y_{\rm O}^{\rm Ia}=0$.

@@

We explore five different functional forms for the SN Ia DTD: a power-law, a broken power-law with an initial flat plateau, an exponential, a two-population model, and a model based on triple-system dynamics, with a one or two variants of each. We present the unnormalized functional forms for $R_{\rm Ia}(t)$ below; VICE automatically normalizes and scales the DTD to produce the correct number of SNe Ia over the lifetime of the Galaxy. Figure \ref{fig:dtds} presents these DTDs as a function of time after star formation, and Table \ref{tab:dtds} summarizes the DTD models and their parameters.

\begin{figure}
    \centering
    \includegraphics{figures/delay_time_distributions.pdf}
    \caption{Selection of models for the SN Ia delay time distribution (DTD) used in this paper. All functions are normalized such that $R_{\rm Ia}(t=1\,{\rm Gyr})=1$. The black squares represent the DTD recovered for the SDSS-II sample of SNe Ia by \citet{Maoz2012-SloanIIDTD} at the same scale as the DTD models. The horizontal and vertical error bars indicate the time range and 1$\sigma$ uncertainties, respectively, of each DTD measurement. The colored circles along the horizontal axis indicate the median delay time for each model.}
    \label{fig:dtds}
    \script{delay_time_distributions.py}
\end{figure}

@@

\paragraph{Power-law} A simple power law with slope $\alpha$:
\begin{equation}
    R_{\rm Ia}(t) \propto (t/1\,\rm{Gyr})^\alpha
    \label{eq:powerlaw-dtd}
\end{equation}
% where $A_\alpha = (\alpha+1)(t_{\rm max}^{\alpha+1} - t_{\rm min}^{\alpha+1})^{-1}$ is the normalization coefficient.
A declining power-law with $\alpha\sim-1$ is expected to arise from typical assumptions about the distribution of post-common envelope separations and the rate of gravitational wave inspiral \citep[see Section 3.5 from][]{Maoz2014-Review}. It is therefore a commonly assumed DTD in GCE studies \citepalias[e.g.,][]{Johnson2021-Migration}. Additionally, the observational evidence for a power-law DTD is strong. \citet{Maoz2017-CosmicDTD} obtain a DTD with $\alpha=-1.07\pm0.09$ based on volumetric rates and an assumed cosmic SFH for field galaxies in redshift range $0\leq z\leq 2.25$. \citet{Wiseman2021-DESRates} obtain a similar slope of $\alpha=-1.13\pm0.05$ for field galaxies in the redshift range $0.2<z<0.6$. For galaxy clusters, \citet{Maoz2017-CosmicDTD} find a steeper DTD slope of $\alpha=-1.39^{+0.32}_{-0.05}$. \citet{Heringer2019-FieldGalaxyDTD} use a SFH-independent method to constrain the DTD for field galaxies within $0.01<z<0.2$ and find a steeper slope of $\alpha=-1.34^{+0.19}_{-0.17}$.
In this paper, we investigate the cases $\alpha=-1.1$ and $\alpha=-1.4$.

\paragraph{Plateau} A modification of the simple power-law in which the DTD ``plateaus'' for a duration $W$ before declining:
\begin{equation}
    R_{\rm Ia}(t) \propto
    \begin{cases}
        1, & t < W \\
        (t/W)^\alpha, & t \ge W
    \end{cases}
    \label{eq:plateau-dtd}
\end{equation}
% where $B_W=[(W-t_{\rm min}) + \frac{t_{\rm max}^{\alpha+1} - W^{\alpha+1}}{(\alpha+1)W^\alpha}]^{-1}$ is the normalization coefficient.
Our primary motivation is to consider a model which matches observations at delay times beyond a few Gyr, where the DTD is best constrained, but which has a much smaller fraction of prompt ($\lesssim 100$ Myr) SNe Ia.
To our knowledge, this form of the DTD has not been considered in previous studies. However, we show in Appendix \ref{app:analytical-dtds} that this form can approximate the more complicated analytical DD DTDs from \citet{Greggio2005-AnalyticalRates}. We investigate the cases $W=0.3$ Gyr and $W=1$ Gyr, taking $\alpha=-1.1$ for all plateau models.

\paragraph{Triple-system} A DTD based on simulations of triple-system evolution by \citet{Rajamuthukumar2023-TripleEvolution}. We approximate their numerically-generated DTD as a special case of the plateau model (Equation \ref{eq:plateau-dtd}) where the initial rate is quite low until an instantaneous rise to the plateau value at time $t_{\rm rise}$:
\begin{equation}
    R_{\rm Ia}(t) \propto
    \begin{cases}
        \epsilon, & t < t_{\rm rise} \\
        1, & t_{\rm rise} \leq t < W \\
        (t / W) ^ \alpha, & t \geq W
    \end{cases}
    \label{eq:triple-dtd}
\end{equation}
with $t_{\rm rise}=0.5$ Gyr, $W=0.5$ Gyr, $\alpha=-1.1$, and $\epsilon=0.05$ so the initial rate is 5\% of the peak rate.

\paragraph{Exponential} An exponentially declining DTD with timescale $\tau$:
\begin{equation}
    R_{\rm Ia}(t) \propto \frac{1}{\tau} e^{-t/\tau}
    \label{eq:exponential-dtd}
\end{equation}
This model can be used to calculate an analytic solution to the equations of chemical equilibrium, making it a popular choice among GCE studies. \citet{Schonrich2009-RadialMixing} and \citet{Weinberg2017-ChemicalEquilibrium} both assume an exponential DTD with a timescale $\tau=1.5$ Gyr. We show in Appendix \ref{app:analytical-dtds} that a $\tau=1.5$ Gyr exponential is an adequate approximation for the analytical SD DTD from \citet{Greggio2005-AnalyticalRates}. 
There is less observational support for an exponential DTD. \citet{Strolger2020-ExponentialDTD}, fitting to the cosmic SFH and SFHs from field galaxies, find a range of exponential-like solutions with timescales $\sim 1.5 - 6$ Gyr.
In this paper we investigate timescales $\tau=1.5$ and 3 Gyr.
% I've looked at Matteucci & Greggio (1986) and I'm not sure what their DTD would be

\paragraph{Two-population} A DTD in which $\sim50\%$ of SNe Ia belong to a ``prompt'' Gaussian component at small $t$ and the remainder form an exponential tail at large $t$:
\begin{equation}
    R_{\rm Ia}(t) \propto 0.5\Big[e^{-\frac{(t-t_p)^2}{2\sigma^2}} + e^{-t/\tau}\Big]
    \label{eq:prompt-dtd}
\end{equation}
To approximate the DTD from \citet{Mannucci2006-TwoPopulations}, we take $t_p=50$ Myr, $\sigma=15$ Myr, and $\tau=3$ Gyr, which results in $\sim 50\%$ of SNe Ia exploding within $t<100$ Myr. This formulation is slightly different than the approximation used in other GCE studies \citep[e.g.,][]{Matteucci2006-BimodalDTDConsequences,Poulhazan2018-PrecisionPollution}, where it has a more distinctly bimodal shape. We have compared the two approximations to this DTD in a one-zone model and found that they produce very similar abundance tracks and MDFs. This DTD is adopted by the Feedback In Realistic Environments \citep[FIRE;][]{Hopkins2014-FIRE-1} and FIRE-2 \citep{Hopkins2018-FIRE-2} simulations.

@@

\subsection{The distribution of [Fe/H]}

>>>>

% user-defined commands
\newcommand{\yes}{\textcolor{green}{\checkmark}}
\newcommand{\meh}{\textcolor{black}{$\sim$}}
\newcommand{\no}{\textcolor{red}{$\times$}}
\newcommand{\osuaffil}{Department of Astronomy, The Ohio State University, 140 W. 18th Ave, Columbus OH 43210, USA}
\newcommand{\ccappaffil}{Center for Cosmology and AstroParticle Physics, The Ohio State University, 191 W. Woodruff Ave., Columbus OH 43210, USA}
\newcommand{\aFe}{[$\alpha$/Fe]\xspace}
\newcommand{\vice}{{\tt VICE}\xspace}

@@

% Abstract from AAS 241
\begin{abstract}
    Type Ia supernovae (SNe Ia) produce most of the Fe-peak elements in the Universe and therefore are a crucial ingredient in galactic chemical evolution models. SNe Ia do not explode immediately after star formation and the delay-time distribution (DTD) has not been definitively determined by supernova surveys or theoretical models. Because the DTD also affects the relationship among age, [Fe/H], and \aFe in chemical evolution models, comparison with observations of stars in the Milky Way is an important consistency check for any proposed DTD. We implement several popular forms of the DTD in combination with multiple star formation histories for the Milky Way in one-zone and multi-zone chemical evolution models using the Versatile Integrator for Chemical Evolution (\vice). \vice includes a prescription for radial stellar migration in multi-zone models. We compare our predicted interstellar medium abundance tracks, stellar abundance distributions, and stellar age distributions to the final data release of the Apache Point Observatory Galactic Evolution Experiment (APOGEE). We find that the SN Ia DTD has the largest effect on the \aFe distribution: a DTD with more prompt SNe Ia produces a stellar abundance distribution that is skewed toward a lower \aFe ratio, whereas a DTD with more delayed SNe Ia produces a distribution that is skewed toward higher \aFe. While the DTD alone cannot explain the observed bimodality in the \aFe distribution, in combination with an appropriate star formation history it affects the goodness of fit between the simulated and observed high-$\alpha$ sequence. Our model results favor a DTD with longer delay times and fewer prompt SNe Ia than the fiducial $t^{-1}$ power law.
\end{abstract}

@@

This paper presents the first comprehensive look at the DTD in a multi-zone GCE model that can qualitatively reproduce the observed abundance structure of the Milky Way disk. A multi-zone approach allows for a radially-dependent parameterization of the SFH, outflows, stellar migration, and abundance gradient which can better match observations across the Galactic disk. We evaluate a selection of DTD models from the literature with multiple SFHs and a prescription for radial stellar migration in the Versatile Integrator for Chemical Evolution \citep[\vice;][]{JohnsonWeinberg2020-Starbursts}. In Section \ref{sec:methods}, we present our models for the DTD and SFH, and we describe our chemical evolution models. In Section \ref{sec:onezone-results}, we present the results of our one-zone chemical evolution models. In Section \ref{sec:multizone-results}, we present the results of our multi-zone models and compare to observations. In Section \ref{sec:discussion}, we discuss the implications for the DTD and future surveys. In Section \ref{sec:conclusions}, we summarize our conclusions.

@@

We make use of the multi-zone simulation tools in \vice developed by \citet{Johnson2021-Migration} (hereafter \citetalias{Johnson2021-Migration}). The basic setup of our models follows theirs: the simulated Galaxy is a disk with a radius of 20 kpc and is divided into concentric rings of width $\delta R_{\rm gal}=100$ pc. The gas in each ring is assumed to be instantaneously mixed; while star particles are allowed to migrate radially (see Appendix \ref{app:migration}), there is no exchange of gas between rings. We run our simulations with a time-step size of $\Delta t=10$ Myr up to a maximum time of $t_{\rm max}=13.2$ Gyr. Following \citetalias{Johnson2021-Migration}, we set \vice to form $n=8$ stellar populations per ring per time-step, and we set a maximum star-formation radius of $R_{\rm SF} = 15.5$ kpc, such that $\dot\Sigma_*=0$ for $R_{\rm gal}>R_{\rm SF}$ (stars are allowed to migrate past $R_{\rm SF}$, they are just not born there). We summarize the parameters for our multi-zone models in Table \ref{tab:multizone-parameters}.

\begin{deluxetable*}{Cccl}
    \tablecaption{A summary of parameters and their fiducial values for our chemical evolution models. We omit parameters that are unchanged from \citetalias{Johnson2021-Migration}; see their Table 1 for details.\label{tab:multizone-parameters}}
    \tablehead{
        \colhead{Quantity} & \colhead{Fiducial Value(s)} & \colhead{Section} & \colhead{Description}
    }
    \startdata
        % Multi-zone simulation parameters
        R_{\rm gal}     & [0, 20] kpc   & \ref{sec:methods} & Galactocentric radius \\
        \delta R_{\rm gal}  & 100 pc    & \ref{sec:methods} & Width of each concentric ring \\
        \Delta R_{\rm gal}  & N/A       & \ref{app:migration} & Change in orbital radius due to stellar migration \\
        p(\Delta R_{\rm gal}|\tau,R_{\rm form}) & Equation \ref{eq:radial-migration}    & \ref{app:migration} & Probability density function of radial migration \\
        z                   & [-3, 3] kpc                & \ref{app:migration} & Distance from Galactic midplane at present day \\
        p(z|\tau,R_{\rm final}) & Equation \ref{eq:sech-pdf}            & \ref{app:migration} & Probability density function of Galactic midplane distance\\
        \Delta t        & 10 Myr    & \ref{sec:methods} & Time-step size \\
        t_{\rm max}     & 13.2 Gyr  & \ref{sec:methods} & Maximum simulation time \\
        n               & 8         & \ref{sec:methods} & Number of stellar populations formed per ring per time-step \\
        R_{\rm SF}      & 15.5 kpc  & \ref{sec:methods} & Maximum radius of star formation \\
        M_{g,0}   & 0         & \ref{sec:sfh}     & Initial gas mass \\
        % DTD
        R_{\rm Ia}(t)   & Equation \ref{eq:dtd-function}    & \ref{sec:dtd-models}  & Delay time distribution of Type Ia supernovae \\
        t_D             & 40 Myr    & \ref{sec:dtd-models}  & Minimum SN Ia delay time \\
        N_{\rm Ia}/M_*  & $2.2\times10^{-3}$ M$_\odot^{-1}$ & \ref{sec:yields}  & SNe Ia per unit mass of stars formed \citep{MaozMannucci2012-SNeIaReview} \\
        % Nucleosynthetic yields
        y_{\rm O}^{\rm CC}  & 0.015     & \ref{sec:yields}  & CCSN yield of O    \\
        y_{\rm Fe}^{\rm CC} & 0.0012    & \ref{sec:yields}  & CCSN yield of Fe   \\
        y_{\rm O}^{\rm Ia}  & 0         & \ref{sec:yields}  & SN Ia yield of O       \\
        y_{\rm Fe}^{\rm Ia} & 0.00214   & \ref{sec:yields}  & SN Ia yield of Fe \\
        % Star formation histories
        f_{\rm IO}(t|R_{\rm gal})   & Equation \ref{eq:insideout-sfh}   & \ref{sec:sfh} & Time-dependence of the inside-out SFR \\
        f_{\rm LB}(t|R_{\rm gal})   & Equation \ref{eq:lateburst-sfh}   & \ref{sec:sfh} & Time-dependence of the late-burst SFR \\
        \uptau_{\rm EB}(t)          & Equation \ref{eq:earlyburst-taustar}  & \ref{sec:sfh}   & Time-dependence of the early-burst SFE timescale \\
        f_{\rm EB}(t|R_{\rm gal})   & Equation \ref{eq:earlyburst-ifr}  & \ref{sec:sfh} & Time-dependence of the early-burst IFR \\
        f_{\rm TI}(t|R_{\rm gal})   & Equation \ref{eq:twoinfall-ifr}   & \ref{sec:sfh} & Time-dependence of the two-infall IFR \\
    \enddata
\end{deluxetable*}

@@

Like \citetalias{Johnson2021-Migration}, we focus our analysis on O and Fe, representing the $\alpha$ and Fe-peak elements, respectively. Both elements are produced by CCSNe. \vice adopts the instantaneous recycling approximation for CCSNe, so the equation which governs CCSN enrichment as a function of star formation for some element $x$ is simply
\begin{equation}
    \dot M_x^{\rm CC} = y_x^{\rm CC} \dot M_*
    \label{eq:ccsn-enrichment}
\end{equation}
where $y_x^{\rm CC}$ is the CCSN yield of element $x$ per unit mass of star formation, and $\dot M_*$ is the SFR. 
Following \citetalias{Johnson2021-Migration}, who in turn adopt their CCSN yields from \citet{ChieffiLimongi2004-CCSNYields} and \citet{LimongiChieffi2006-CCSNYields}, we adopt $y_{\rm O}^{\rm CC}=0.015$ and $y_{\rm Fe}^{\rm CC}=0.0012$. The primary effect of these yields is to set the low-[Fe/H] ``plateau'' in [O/Fe] which represents pure CCSN enrichment. The chosen yields for this paper produce a plateau at ${\rm [O/Fe]}=0.45$; see \citet{Weinberg2023-CCSNYield} for more discussion on the effect of the CCSN yields on chemical evolution.

@@

The yield $y_{\rm Fe}^{\rm Ia}$ measures the mass of Fe produced by SNe Ia per unit mass of star formation. It can be expressed in terms of the integral of the DTD:
\begin{equation}
    y_{\rm Fe}^{\rm Ia} = m_{\rm Fe}^{\rm Ia} \int_{t_D}^{t_{\rm max}} R_{\rm Ia}(t')dt' = m_{\rm Fe}^{\rm Ia} \frac{N_{\rm Ia}}{M_*},
    \label{eq:dtd-integral}
\end{equation}
where $m_{\rm Fe}^{\rm Ia}$ is the average mass of Fe produced by a single SN Ia, and $N_{\rm Ia}/M_*=2.2\times10^{-3}\,{\rm M}_\odot^{-1}$ is the average number of SNe Ia per mass of stars formed \citep{MaozMannucci2012-SNeIaReview}. 
Adjusting the value of $y_{\rm Fe}^{\rm Ia}$ primarily affects the end point of chemical evolution tracks in [O/Fe]--[Fe/H] space. 
Following \citetalias{Johnson2021-Migration}, we adopt $y_{\rm Fe}^{\rm Ia}=0.00214$. This yield was originally adapted from the W70 model of \citet{Iwamoto1999-SNIaYields}, but was slightly modified so that the inside-out SFH model produces stars with ${\rm [O/Fe]}\approx 0.0$ by the end of the simulation.
\citet{Palla2021-SNIaYield} studies the effect of different SN Ia yields on GCE models in detail.

@@

We explore five different functional forms for the SN Ia DTD: a power-law, a broken power-law with an initial flat plateau, an exponential, a two-population model, and a model based on triple-system dynamics, with a one or two variants of each. Figure \ref{fig:dtds} presents these DTDs as a function of time after star formation, and Table \ref{tab:dtds} summarizes the DTD models and their parameters. We use simple forms rather than simulated physical or analytical models of SNe Ia for the sake of decreased computational time. Additionally, physically-motivated models of the DTD must contend with many unknown or poorly-constrained parameters, so our simple forms allow us to reduce the number of free parameters. In Appendix \ref{app:analytical-dtds}, we show that a few of our simple forms adequately approximate the more complete analytical models of \citet{Greggio2005-AnalyticalRates}.

In this subsection we present unnormalized functional forms $f_{\rm Ia}$ for each DTD model. The relationship between $f_{\rm Ia}$, which has units of ${\rm Gyr}^{-1}$, and $R_{\rm Ia}$, which has units of ${\rm M}_\odot^{-1}\,{\rm Gyr}^{-1}$, is: 
\begin{equation}
    R_{\rm Ia}(t) = 
    \begin{cases}
        \frac{N_{\rm Ia}}{M_*}
        \frac{f_{\rm Ia}(t)}{\int_{t_D}^{t_{\rm max}} f_{\rm Ia}(t') dt'}, & t \ge t_D \\
        0 & t < t_D.
    \end{cases}
    \label{eq:dtd-function}
\end{equation}
\vice automatically normalizes the provided function for the DTD according to this equation.

\begin{figure}
    \centering
    \includegraphics{figures/delay_time_distributions.pdf}
    \caption{Selection of models for the SN Ia delay time distribution (DTD) used in this paper. All functions are normalized such that $f_{\rm Ia}(t=1\,{\rm Gyr})=1$. The black squares represent the DTD recovered for the SDSS-II sample of SNe Ia by \citet{Maoz2012-SloanIIDTD} at the same scale as the DTD models. The horizontal and vertical error bars indicate the time range and 1$\sigma$ uncertainties, respectively, of each DTD measurement. The colored circles along the horizontal axis indicate the median delay time for each model.}
    \label{fig:dtds}
    \script{delay_time_distributions.py}
\end{figure}

@@

\paragraph{Power-law} A simple power law with slope $\alpha$:
\begin{equation}
    f_{\rm Ia}^{\rm plaw}(t) = (t/1\,\rm{Gyr})^\alpha
    \label{eq:powerlaw-dtd}
\end{equation}
% where $A_\alpha = (\alpha+1)(t_{\rm max}^{\alpha+1} - t_{\rm min}^{\alpha+1})^{-1}$ is the normalization coefficient.
A declining power-law with $\alpha\sim-1$ is expected to arise from typical assumptions about the distribution of post-common envelope separations and the rate of gravitational wave inspiral \citep[see Section 3.5 from][]{Maoz2014-Review}. It is therefore a commonly assumed DTD in GCE studies (e.g., \citealt{Rybizki2017-Chempy}; \citetalias{Johnson2021-Migration}). Additionally, the observational evidence for a power-law DTD is strong. \citet{Maoz2017-CosmicDTD} obtain a DTD with $\alpha=-1.07\pm0.09$ based on volumetric rates and an assumed cosmic SFH for field galaxies in redshift range $0\leq z\leq 2.25$. \citet{Wiseman2021-DESRates} obtain a similar slope of $\alpha=-1.13\pm0.05$ for field galaxies in the redshift range $0.2<z<0.6$. For galaxy clusters, \citet{Maoz2017-CosmicDTD} find a steeper DTD slope of $\alpha=-1.39^{+0.32}_{-0.05}$. \citet{Heringer2019-FieldGalaxyDTD} use a SFH-independent method to constrain the DTD for field galaxies within $0.01<z<0.2$ and find a steeper slope of $\alpha=-1.34^{+0.19}_{-0.17}$.
In this paper, we investigate the cases $\alpha=-1.1$ and $\alpha=-1.4$.

\paragraph{Plateau} A modification of the simple power-law in which the DTD ``plateaus'' for a duration $W$ before declining:
\begin{equation}
    f_{\rm Ia}^{\rm plat}(t) =
    \begin{cases}
        1, & t < W \\
        (t/W)^\alpha, & t \ge W
    \end{cases}
    \label{eq:plateau-dtd}
\end{equation}
% where $B_W=[(W-t_{\rm min}) + \frac{t_{\rm max}^{\alpha+1} - W^{\alpha+1}}{(\alpha+1)W^\alpha}]^{-1}$ is the normalization coefficient.
Our primary motivation is to consider a model which matches observations at delay times beyond a few Gyr, where the DTD is best constrained, but which has a much smaller fraction of prompt ($\lesssim 100$ Myr) SNe Ia than the simple power law.
To our knowledge, this form of the DTD has not been considered in previous studies. However, we show in Appendix \ref{app:analytical-dtds} that this form can approximate the more complicated analytical DD DTDs from \citet{Greggio2005-AnalyticalRates}. We investigate the cases $W=0.3$ Gyr and $W=1$ Gyr, taking $\alpha=-1.1$ for all plateau models.

\paragraph{Triple-system} A DTD based on simulations of triple-system evolution by \citet{Rajamuthukumar2023-TripleEvolution}. We approximate their numerically-generated DTD as a special case of the plateau model (Equation \ref{eq:plateau-dtd}) where the initial rate is quite low until an instantaneous rise to the plateau value at time $t_{\rm rise}$:
\begin{equation}
    f_{\rm Ia}^{\rm triple}(t) =
    \begin{cases}
        \epsilon, & t < t_{\rm rise} \\
        1, & t_{\rm rise} \leq t < W \\
        (t / W) ^ \alpha, & t \geq W
    \end{cases}
    \label{eq:triple-dtd}
\end{equation}
with $t_{\rm rise}=0.5$ Gyr, $W=0.5$ Gyr, $\alpha=-1.1$, and $\epsilon=0.05$ so the initial rate is 5\% of the peak rate.

\paragraph{Exponential} An exponentially declining DTD with timescale $\tau$:
\begin{equation}
    f_{\rm Ia}^{\rm exp}(t) = \frac{1}{\tau} e^{-t/\tau}
    \label{eq:exponential-dtd}
\end{equation}
This model can be used to calculate an analytic solution to the equations of chemical equilibrium, making it a popular choice among GCE studies. \citet{Schonrich2009-RadialMixing} and \citet{Weinberg2017-ChemicalEquilibrium} both assume an exponential DTD with a timescale $\tau=1.5$ Gyr. We show in Appendix \ref{app:analytical-dtds} that a $\tau=1.5$ Gyr exponential is an adequate approximation for the analytical SD DTD from \citet{Greggio2005-AnalyticalRates}. 
There is less observational support for an exponential DTD. \citet{Strolger2020-ExponentialDTD}, fitting to the cosmic SFH and SFHs from field galaxies, find a range of exponential-like solutions with timescales $\sim 1.5 - 6$ Gyr.
In this paper we investigate timescales $\tau=1.5$ and 3 Gyr.
% I've looked at Matteucci & Greggio (1986) and I'm not sure what their DTD would be

\paragraph{Two-population} A DTD in which $\sim50\%$ of SNe Ia belong to a ``prompt'' Gaussian component at small $t$ and the remainder form an exponential tail at large $t$:
\begin{equation}
    f_{\rm Ia}^{\rm twopop}(t) = 0.5\Big[e^{-\frac{(t-t_p)^2}{2\sigma^2}} + e^{-t/\tau}\Big]
    \label{eq:prompt-dtd}
\end{equation}
To approximate the DTD from \citet{Mannucci2006-TwoPopulations}, we take $t_p=50$ Myr, $\sigma=15$ Myr, and $\tau=3$ Gyr, which results in $\sim 50\%$ of SNe Ia exploding within $t<100$ Myr. This formulation is slightly different than the approximation used in other GCE studies \citep[e.g.,][]{Matteucci2006-BimodalDTDConsequences,Poulhazan2018-PrecisionPollution}, where it has a more distinctly bimodal shape. We have compared the two approximations to this DTD in a one-zone model and found that they produce very similar abundance tracks and MDFs. This DTD is adopted by the Feedback In Realistic Environments \citep[FIRE;][]{Hopkins2014-FIRE-1} and FIRE-2 \citep{Hopkins2018-FIRE-2} simulations.

@@

\subsection{The distribution of [Fe/H]}
\label{sec:feh-df}
